### 2016-05-12 refs/ReviewPapers.tex
papers review version 1

\documentclass{article}
\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{2.0in}
%\addtolength{\textwidth}{1.75in}
\addtolength{\topmargin}{-1.0in}
%\addtolength{\textheight}{1.80in}
\addtolength{\textheight}{1.95in}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{sidecap}
\usepackage[pdftex]{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\decimalpoint

\begin{document}
\section{Evidencias observacionales de planos:}
\noindent The VPOS (Pawlowski et al 2012: https://arxiv.org/abs/1204.5176)\\
A vast thin plane of corrotating dwarf galaxies orbiting the Andromeda
Galaxy (Ibata et al 2013)\\
Corotation in SDSS data (Ibata et al 2014)\\
There was a reply to this paper (A new spin on disks of satellite
galaxies, Cuatun et al 2014)\\
Planes in Centaurus A (Tully et al 2015)\\
Varios papers de pawlowski que muestran que los nuevos sat\'elites de
MW estan en el plano, que el plano no se debe al Sdss footprint, que
los globular clusters tambien estan en el plano... etc,\\ 
Papers de Conn donde encuentra el plano\\

\section{B\'usqueda de planos en simulaciones:}
\noindent A comparison of the distribution of satellite galaxies
around Andromeda and the results of $\Lambda$CDM simulations (Bahl and
Baumgardt 2013) Encuentran planos en Millenium II.\\
A thousand shadows of Andromeda: rotating planes of satellites in the
Millennium-II cosmological simulation (Ibata et al 2013) dicen que el
paper de Bahl y Baumgardt est\'a mal y que no hay planos en millenium
II\\
Co-orbiting satellite galaxy structures are still in conflict with the
distribution of primordial dwarf galaxies (Pawlowski et al 2014)\\
Co-orbiting planes of sub-halos are similarly unlikely around paired
and isolated hosts (Pawlowski et al 2014)\\
Vast planes of satellites in a high resolution simulation of the Local
Group: comparison to Andromeda (Gillet et al 2014). finds a plnes
similar to that of M31 in Clues simulations.\\
Planes of satellite galaxies: when exceptions are the rule (Cautun et
al 2015) encuentra que 10 por ciento de los halos tienen planos
iguales o m\'as prominentes que los del LG.\\

\section{Posibles or\'igenes de los planos:}
\noindent Preferential accretion (Libeskind et al 20??)\\
Alignments with the cosmic web (Tempel et al 20??)\\
The vast thin plane of M31 co-rotating dwarfs: an additional fossil
signature of the M31 merger and of its considerable impact in the
whole Local Group (Hammer et al 2013) Major merger in M31-MW system
plane galaxies are tidal dwarfs (n-body simulations)\\
Kroupa tiene todo un carretazo de que las dwarfs son todas tidal
dwarfs\\
MIRAR lo que est\'a haciendo Pierre Alan Duc con tidal dwrafs porque en
una conferencia este a;o habl\'o de un posible escenario
intermedio...\\
Evidence for Early Filamentary Accretion from the Andromeda Galaxy's
Thin Plane of Satellites (Buck et al 2015)\\
Alignments between galaxies, satellite systems and haloes (Cautun et al
2016) NO LO HE LEIDO...

\section{Problemas con esas explicaciones:}
The Vast Polar Structure of the Milky Way and Filamentary Accretion of
Sub-Halos (Pawlowski et al 2012)\\
Paper de Collins (2013 y 2016) donde explica que no hay doferencias en
las propiedades de los on-plane y los off-plane satellites\\
Problemas con la estabilidad a largo plazo de los planos: Bowden et al
2012, Gonzales et al 2015...\\
The Plane Truth: Andromeda analog thin Planes of Satellites are not
kinematical coherent structures (Buck et al 2015)\\

\end{document}